Cooperative Kernels: GPU Multitasking for Blocking Algorithms (Extended Version)

### 2017-07-06 write-ups/FSE/extended/paper.tex
Extended version of paper.

%%% The following is specific to ESEC/FSE'17 and the paper
%%% 'Cooperative Kernels: GPU Multitasking for Blocking Algorithms'
%%% by Tyler Sorensen, Hugues Evrard, and Alastair F. Donaldson.
%%%
\setcopyright{acmlicensed}
\acmDOI{10.1145/3106237.3106265}
\acmISBN{978-1-4503-5105-8/17/09}
\acmConference[ESEC/FSE'17]{2017 11th Joint Meeting of the European Software Engineering Conference and the ACM SIGSOFT Symposium on the Foundations of Software Engineering}{September 4--8, 2017}{Paderborn, Germany}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}

@@

\title{Cooperative Kernels: GPU Multitasking for Blocking Algorithms}

@@

We present our cooperative kernels programming model as an extension
to OpenCL.
%However, the cooperative kernels concept is more general,
%and could be applied to extend other GPU programming models, such as
%CUDA and HSA.
We describe the semantics of the model
(\mysec\ref{sec:semantics}), use our motivating examples to discuss
programmability (\mysec\ref{sec:programmingguidelines}) and outline
important nonfunctional properties that the model requires to work
successfully (\mysec\ref{sec:nonfunctional}).
%and explain that the
%model is backwards-compatible
%(\mysec\ref{sec:backwardscompatibility}).
%

@@

%Among the GPUs available to us, four met our requirements: Intel HD
%520, HD 5500 and Iris 6100, and AMD Radeon R7.  However, AMD's Linux
%drivers for OpenCL suffer from a known defect whereby long-running
%kernels lead to unkillable defunct
%processes~\cite{amd-defunct-process}.
%
%and simultaneously obtained alarming results on this platform when
%using SVM atomics: modifying a kernel to include a simple host/device
%handshake through SVM slowed down kernel execution by an order of
%magnitude in some cases.  This behaviour, combined with the defunct
%processes issue, lead us to suspect that the support we require for
%cooperative kernels is not yet mature enough within AMD's drivers,
%which we plan to discuss with engineers at AMD.
%
\TSAdded{We thus ran our experiments on three Intel GPUs: HD 520, HD 5500
and Iris 6100. The results were similar across the GPUs, so for
conciseness, we report only on the Iris 6100 GPU (driver
20.19.15.4463) with a host CPU i3-5157U. The Iris has a reported 47
compute units. Results for the other Intel GPUs are available in
supplementary material, found on the author's website.}

@@

\myparagraph{Responsiveness}
%
\myfiglong\ref{fig:fine-grained-timing} reports, on three
configurations, the average gather and execution times for the
non-cooperative kernel with respect to the quantity of workgroups allocated to
it.  A logarithmic scale is used for time since gather times tend to
be much smaller than execution times. The horizontal grey lines
indicates the desired period for non-cooperative kernels.  These
graphs show a representative sample of our results; the full set of
graphs for all configurations is provided in supplementary material.

@@

\iffalse

@@

\fi
\end{document}

>>>>

%%% The following is specific to ESEC/FSE'17 and the paper
%%% 'Cooperative Kernels: GPU Multitasking for Blocking Algorithms'
%%% by Tyler Sorensen, Hugues Evrard, and Alastair F. Donaldson.
%%%
\setcopyright{acmlicensed}
%\acmDOI{10.1145/3106237.3106265}
%\acmISBN{978-1-4503-5105-8/17/09}
%\acmConference[ESEC/FSE'17]{2017 11th Joint Meeting of the European Software Engineering Conference and the ACM SIGSOFT Symposium on the Foundations of Software Engineering}{September 4--8, 2017}{Paderborn, Germany}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}

@@

\title{Cooperative Kernels: GPU Multitasking for Blocking Algorithms (Extended Version)}

@@

We present our cooperative kernels programming model as an extension
to OpenCL.
%However, the cooperative kernels concept is more general,
%and could be applied to extend other GPU programming models, such as
%CUDA and HSA.
We describe the semantics of the model
(\mysec\ref{sec:semantics}), presenting a more formal operational semantics in Appendix~\ref{appendix:semantics}
and discussing possible alternative semantic choices in Appendix~\ref{appendix:semanticalternatives}, use our motivating examples to discuss
programmability (\mysec\ref{sec:programmingguidelines}) and outline
important nonfunctional properties that the model requires to work
successfully (\mysec\ref{sec:nonfunctional}).
%and explain that the
%model is backwards-compatible
%(\mysec\ref{sec:backwardscompatibility}).
%

@@

%Among the GPUs available to us, four met our requirements: Intel HD
%520, HD 5500 and Iris 6100, and AMD Radeon R7.  However, AMD's Linux
%drivers for OpenCL suffer from a known defect whereby long-running
%kernels lead to unkillable defunct
%processes~\cite{amd-defunct-process}.
%
%and simultaneously obtained alarming results on this platform when
%using SVM atomics: modifying a kernel to include a simple host/device
%handshake through SVM slowed down kernel execution by an order of
%magnitude in some cases.  This behaviour, combined with the defunct
%processes issue, lead us to suspect that the support we require for
%cooperative kernels is not yet mature enough within AMD's drivers,
%which we plan to discuss with engineers at AMD.
%
\TSAdded{We thus ran our experiments on three Intel GPUs: HD 520, HD 5500
and Iris 6100. The results were similar across the GPUs, so for
conciseness, we report only on the Iris 6100 GPU (driver
20.19.15.4463) with a host CPU i3-5157U. The Iris has a reported 47
compute units. Results for the other Intel GPUs are presented in Appendix~\ref{appendix:extragraphs}.}

@@

\myparagraph{Responsiveness}
%
\myfiglong\ref{fig:fine-grained-timing} reports, on three
configurations, the average gather and execution times for the
non-cooperative kernel with respect to the quantity of workgroups allocated to
it.  A logarithmic scale is used for time since gather times tend to
be much smaller than execution times. The horizontal grey lines
indicates the desired period for non-cooperative kernels.  These
graphs show a representative sample of our results; the full set of
graphs for all configurations is provided in Appendix~\ref{appendix:extragraphs}.

@@

@@

\end{document}
